\textbf{ZKLoRA: Efficient Zero-Knowledge Proofs for LoRA Verification}

### 2025-01-16 paper/main.tex
add bagel logo

\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, positioning, calc, fit}

\title{\textbf{ZKLoRA: Efficient Verification of LoRA-Base Model Compatibility for Large Language Models}}
\author{
    \textbf{Bidhan Roy, Peter Potash, Marcos Villagra} \\
    Bagel Research Team\footnotemark[1] \\
    \texttt{bidhan@bagel.net, peter@bagel.net, marcos@bagel.net}
}
\date{\today}

@@

\begin{abstract}
    Parameter-efficient fine-tuning (PEFT) techniques, including Low-Rank Adaptation (LoRA), are widely adopted to customize large-scale language models in distributed training environments. A base model owner may outsource LoRA development to an external contributor, creating two requirements: the owner must verify that the LoRA update stems from the designated open-source model, and the contributor must protect proprietary parameters until compensation is assured.

@@

Figures~\ref{fig:proof_vs_numlora} and \ref{fig:settings_vs_numlora} similarly show that \emph{proof generation} and \emph{settings time} scale with the number of LoRA modules, but also depend heavily on each module’s average size (indicated by larger or smaller dot size). These two steps (proof generation on the LoRA Owner’s side and cryptographic circuit setup) can become more expensive, yet remain feasible in decentralized settings or paid contract relationships. The \emph{Base Model User}, meanwhile, benefits from the relatively short verification overhead.

@@

A typical workflow involves:
\begin{enumerate}
    \item \textbf{Base Model User} possessing a large LLM. 
    \item \textbf{LoRA Owner} having proprietary low-rank adapter weights. 
\end{enumerate}
However, these two parties must confirm the LoRA truly “fits” the base model. Using \texttt{ZKLoRA}, the LoRA Owner proves the adapter’s compatibility \emph{without} exposing private weights:

@@

    \item \textbf{LoRA Transform + Proof Gen.}
    The LoRA Owner applies private LoRA updates, constructing a zero-knowledge proof to attest correct usage.

@@

    \item \textbf{Evaluate.}
    With trust in the LoRA’s correctness, the user measures performance (e.g., loss).
\end{enumerate}

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
    font=\small,
    >=latex,
    thick,
    block/.style={
      draw,
      rounded corners,
      align=center,
      minimum width=1.65cm,  % Decreased from 1.8cm to 1.65cm
      minimum height=0.9cm,
      fill=gray!10  % Light gray background for regular blocks
    },
    dashedblock/.style={
      draw,
      dashed,
      rounded corners,
      align=center,
      font=\footnotesize,
      fill=orange!10,  % Light orange background for LoRA block
      minimum width=1.65cm,  % Decreased to match other blocks
      minimum height=0.9cm
    },
    arrow/.style={->, thick},
    node distance=1.65cm  % Decreased from 1.8cm to 1.65cm
]
\node[block, label=above:\textcolor{teal}{\textbf{Input}}] (input) {$\mathbf{x}$};
\node[block, right=1.65cm of input] (W1) {$\mathbf{W}_1$};
\node[dashedblock, above=1.6cm of W1, label=above:\textcolor{orange!70!black}{\textbf{LoRA Owner}}] (LoRA) {LoRA for Layer 1};
\node[block, right=1.65cm of W1] (act) {ReLU};
\node[block, right=1.65cm of act] (W2) {$\mathbf{W}_2$};
\node[block, label=above:\textcolor{teal}{\textbf{Output}}, right=1.65cm of W2] (output) {$\mathbf{y}$};
\draw[->, teal!60!black] (input) -- node[above, font=\scriptsize]{local multiply} (W1);
\draw[->, dashed, orange!70!black] (W1.north) to[bend left=15] node[left, font=\scriptsize]{partial acts} (LoRA.south);
\draw[->, dashed, orange!70!black] (LoRA.south) to[bend left=15] node[right, font=\scriptsize]{updated $\mathbf{z}'$} (W1.north);
\draw[->, teal!60!black] (W1) -- node[above, font=\scriptsize]{$\mathbf{z}'$} (act);
\draw[->, teal!60!black] (act) -- node[above, font=\scriptsize]{hidden} (W2);
\draw[->, teal!60!black] (W2) -- (output);
\node[draw, rounded corners, thick, color=blue!40!black,
      fit=(input)(W1)(act)(W2)(output),
      label={[font=\small, color=blue!40!black, xshift=2cm]above:\textbf{Base Model User}},
      inner sep=0.6cm
] (baseModelFit) {};
\end{tikzpicture}
\caption{zklora inference flow on a 2-layer neural network where the first layer has private LoRA weights added.}
\label{fig:toy-mlp-lora}
\end{figure}

@@

\paragraph{(3) Witness Creation.}
The LoRA Owner runs the partial activations through the circuit, tracking all wire values in a “witness.”

\paragraph{(4) Proof.}
Using the witness plus proving key, the LoRA Owner constructs a proof.

@@

\texttt{ZKLoRA} provides a fast, robust mechanism to ensure that private LoRA modules remain consistent with a large base model. Our evaluations indicate that each LoRA module’s correctness can be verified in about 1--2 seconds, even for multi-billion-parameter LLMs. This efficiency bridges the gap between privacy-preserving LoRA development and practical, real-time validation in large-scale deployments. Future expansions could integrate multi-owner LoRAs, advanced zero-knowledge proofs for further performance gains, or partial data-privacy frameworks to shield user inputs as well as LoRA parameters.

>>>>

\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage[export]{adjustbox}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, positioning, calc, fit}

\title{\textbf{ZKLoRA: Efficient Zero-Knowledge Proofs for LoRA Verification}}
\author{ \includegraphics[height=8em]{figs/bagel-logo-bw.png} \\
    \textbf{Bidhan Roy, Peter Potash, Marcos Villagra} \\
    Bagel Research Team\footnotemark[1] \\
    \texttt{bidhan@bagel.net, peter@bagel.net, marcos@bagel.net}
}
\date{\today}

@@

\begin{abstract}
    Parameter-efficient fine-tuning (PEFT) techniques, including Low-Rank Adaptation (LoRA), are widely adopted to customize large-scale language models in distributed training environments. A base model user may want to use LoRA weights developed by an external contributor, creating two requirements: the base model user must verify that the LoRA weights are effective when combined with the designated open-source model, and the contributor must protect proprietary parameters until compensation is assured.

@@

Figures~\ref{fig:proof_vs_numlora} and \ref{fig:settings_vs_numlora} similarly show that \emph{proof generation} and \emph{settings time} scale with the number of LoRA modules, but also depend heavily on each module’s average size (indicated by larger or smaller dot size). These two steps (proof generation on the LoRA contributor’s side and cryptographic circuit setup) can become more expensive, yet remain feasible in decentralized settings or paid contract relationships. The \emph{Base Model User}, meanwhile, benefits from the relatively short verification overhead.

@@

A typical workflow involves:
\begin{enumerate}
    \item \textbf{Base Model User} possessing a large LLM. 
    \item \textbf{LoRA Contributor} having proprietary low-rank adapter weights. 
\end{enumerate}
However, these two parties must confirm the LoRA truly “fits” the base model. Using \texttt{ZKLoRA}, the LoRA Contributor proves the adapter’s compatibility \emph{without} exposing private weights:

@@

    \item \textbf{LoRA Transform + Proof Gen.}
    The LoRA Contributor applies private LoRA forward passes based on partial activations, constructing a zero-knowledge proof to attest correct usage.

@@

    \item \textbf{Evaluate.}
    With trust in the LoRA’s correctness, the user measures performance (e.g. loss).
\end{enumerate}

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
    font=\small,
    >=latex,
    thick,
    block/.style={
      draw,
      rounded corners,
      align=center,
      minimum width=1.65cm,  % Decreased from 1.8cm to 1.65cm
      minimum height=0.9cm,
      fill=gray!10  % Light gray background for regular blocks
    },
    dashedblock/.style={
      draw,
      dashed,
      rounded corners,
      align=center,
      font=\footnotesize,
      fill=orange!10,  % Light orange background for LoRA block
      minimum width=1.65cm,  % Decreased to match other blocks
      minimum height=0.9cm
    },
    arrow/.style={->, thick},
    node distance=1.65cm  % Decreased from 1.8cm to 1.65cm
]
\node[block, label=above:\textcolor{teal}{\textbf{Input}}] (input) {$\mathbf{x}$};
\node[block, right=1.65cm of input] (W1) {$\mathbf{W}_1$};
\node[dashedblock, above=1.6cm of W1, label=above:\textcolor{orange!70!black}{\textbf{LoRA Contributor}}] (LoRA) {LoRA for Layer 1};
\node[block, right=1.65cm of W1] (act) {ReLU};
\node[block, right=1.65cm of act] (W2) {$\mathbf{W}_2$};
\node[block, label=above:\textcolor{teal}{\textbf{Output}}, right=1.65cm of W2] (output) {$\mathbf{y}$};
\draw[->, teal!60!black] (input) -- node[above, font=\scriptsize]{local multiply} (W1);
\draw[->, dashed, orange!70!black] (W1.north) to[bend left=15] node[left, font=\scriptsize]{partial acts} (LoRA.south);
\draw[->, dashed, orange!70!black] (LoRA.south) to[bend left=15] node[right, font=\scriptsize]{updated $\mathbf{z}'$} (W1.north);
\draw[->, teal!60!black] (W1) -- node[above, font=\scriptsize]{$\mathbf{z}'$} (act);
\draw[->, teal!60!black] (act) -- node[above, font=\scriptsize]{hidden} (W2);
\draw[->, teal!60!black] (W2) -- (output);
\node[draw, rounded corners, thick, color=blue!40!black,
      fit=(input)(W1)(act)(W2)(output),
      label={[font=\small, color=blue!40!black, xshift=2cm]above:\textbf{Base Model User}},
      inner sep=0.6cm
] (baseModelFit) {};
\end{tikzpicture}
\caption{zklora inference flow on a 2-layer neural network where the first layer has private LoRA weights added.}
\label{fig:toy-mlp-lora}
\end{figure}

@@

\paragraph{(3) Witness Creation.}
The LoRA Contributor runs the partial activations through the circuit, tracking all wire values in a “witness.”

\paragraph{(4) Proof.}
Using the witness plus proving key, the LoRA Contributor constructs a proof.

@@

\texttt{ZKLoRA} provides a fast, robust mechanism to ensure that private LoRA modules remain consistent with a large base model. Our evaluations indicate that each LoRA module’s correctness can be verified in about 1--2 seconds, even for multi-billion-parameter LLMs. This efficiency bridges the gap between privacy-preserving LoRA development and practical, real-time validation in large-scale deployments. Future expansions could integrate multi-contributor LoRAs, advanced zero-knowledge proofs for further performance gains, or partial data-privacy frameworks to shield user inputs as well as LoRA parameters.
